\chapter*{Introduction}

In 1951 the Belousov-Zhabotinsky class of chemical reactions were created for the first time in a laboratory. This feat marked an
upheaval in thermodynamics, which had now been shown, experimentally, to be ruled by non-equilibrium dynamics even for long mixing
times. This generated for the first time interest in the field of coupled oscillators by the broader scientific community. In the
decades that followed, other areas of research independently led to the same interests. In 1958, Norbet Weiner at the MIT was looking
at power spectrum graphs generated from brain scans and decided that there must be some kind of synchronization amongst coupled,
periodically-firing, neurons. Also inspired by many biological rhythms observed in nature, Arthur Winfree published in 1967 a paper on
the mathematical modelling of interacting oscillators. Winfree defined a phase oscillator as an abstraction of any system that
possesses periodicity. As such, a complicated periodic process inside a living cell could be described by a single real number, between
$0$ and $2\pi$, and its rate of change. The major simplification comes from the fact that when two such phase oscillators interact,
each influences \textit{only} the rate of change in phase of the other. Their trajectories in phase space are assumed to remain closed
loops, and thus the only effective change can be modeled as a change in speed at which they traverse it.

Seven years after Winfree's paper from 1967, the then young Japanese scientist Yoshiki Kuramoto recognized what made Winfree's model
hard to solve. He proposed a modified version of it, assuming a particular form for the coupling function between oscillators. This led
to an arguably less realistic model, but one which Kuramoto was cunningly able to solve. This modified version, together with its
solution, would later be recognized as the Kuramoto model, becoming one of the major founding stones in the field.

Since then, a plethora of novel coupled oscillator models have been put forward. A class of them deal with discrete phase oscillators,
where the phase value that describes a unit can only attain values in a finite set of possible phase states. Discrete phase models are
used to describe systems which present markedly distinct states. An example of such a system is interacting neurons, where one could
identify three distinct phases of a neuron cell as the three available phase values: firing, refractory, ready-to-fire.

Another element present in some models is the introduction of stochasticity in the dynamics, which are deterministic in the Kuramoto
model. The element of stochasticity may be included to account for noise, such as in models that are not closed systems and interact
with some sort of heat bath.

Here we consider discrete phase models with three possible phase states, and stochasticity plays a role in the rate of change of a
units phase state, being the probability of ``jumping'' to the next state per unit time. This model was first introduced by Kevin Wood,
C. Van den Broeck, R. Kawai and Katja Lindenberg \cite{Wood06a}, with the main motivation of finding the simplest possible dynamics
between phase oscillators that still led to a phase transition to synchrony. Initial investigations focused mainly on square lattices
and all-to-all connections, exposing the ubiquity of this type of self-organization. Often times, networks of coupled oscillators
observed in nature belong to neither of these two classes. Neurons, fireflies, cardiac cells, are all examples of systems that do not
operate on lattices or complete-graph networks. One common type of interaction network observed in real systems are a mix between local
and global interactions. Various algorithms have been proposed for the generation of such intermediate connectivity graphs, some of
which generate small-world networks, characterized by the presence of long range interactions while preserving strong local clustering.

A particular example for the generation of small-world networks is the Watts-Strogatz algorithm, which randomly breaks connections on
an existing graph and re-creates them between random vertices. When applied to certain starting graphs with strong local clustering,
the result is a small-world network. Here we investigate how the dynamics of the three-phase oscillators model unfold on such networks,
via simulation and a mean-field analysis.
